---
title: "California Oil Spills"
author: "Bri Baker"
date: "2/18/2021"
output:
  html_document:
    theme: cosmo
    code_folding: hide
    highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(tmap)
library(sf)

```

## Summary

The following are visualizations of oil spill incidents in the State of California in 2008 as tracked by the Office of Spill Prevention and Response (OSPR) Incident Tracking Database. In this database, an "incident" is defined as "a discharge or threatened discharge of petroleum or other deleterious material into the waters of the state" (Lampinen and Conway, 2009).



```{r data}

# read in CA oil spill data
ca_oil <- read_sf(here("part1_spatial/Oil_Spill_Incident_Tracking"), layer = "Oil_Spill_Incident_Tracking") %>% 
  clean_names()

#st_crs(ca_oil)

# Read in the CA county data (TIGER shapefile):
ca_counties <- read_sf(here("part1_spatial/ca_counties"), layer = "CA_Counties_TIGER2016") %>% 
  clean_names() %>% 
  select(name)

#st_crs(ca_counties)

```


## Interactive Map

This interactive map shows both marine and inland oil spill events for California in 2008 (Figure 1).

```{r interactive}

tmap_mode("view") # interactive mode

tm_shape(ca_oil)+
  tm_dots()

```
Figure 1. Interactive map of 2008 California oil spill incident locations.



## Inland oil spill events

The following chloropleth shows only inland oil spill incidents (Figure 2).

```{r inland}
inland_spill <- ca_oil %>%
  filter(inlandmari == "Inland")

spill_county <- ca_counties %>% 
  st_join(inland_spill) %>% 
  count(name)

```

```{r chloropleth}
ggplot(data = spill_county) +
  geom_sf(aes(fill = n), color = "white", size = 0.1) +
  scale_fill_gradientn(colors = c("lightgray","orange","red")) +
  theme_minimal() +
  labs(fill = "Oil spill frequency",
       title = "Inland oil spill frequency per California county in 2008",
       caption = "Bri Baker, 2021 \nData: Lampinen and Conway, 2009") +
  theme(legend.position = c(0.75, 0.8), # move legend
        legend.background = element_rect(fill="white", 
                                         linetype = "solid", color = "whitesmoke"))
```

Figure 2. Chloropleth of inland oil spill frequency per California county in 2008. Colors range from red (high frequency of spills) to grey (low frequency of spills).



## Citation
Lampinen, M. & Conway, C. Oil Spill Incident Tracking [ds394] GIS Dataset. https://map.dfg.ca.gov/metadata/ds0394.html (2009).

